---
title: 'OpenBUGS in R'
date: 'Rendered: `r format(Sys.time(), "%b %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## OpenBUGS 

This is an example of running [`R2OpenBUGS`](https://cran.r-project.org/web/packages/R2OpenBUGS/index.html) inside this project's `docker` container. For more information on `Bayesian inference Using Gibbs Sampling (BUGS)`, please visit [this page](http://www.openbugs.net/w/FrontPage). 

## Example 

### Library

```{r library}
library(R2OpenBUGS)
```

### Data 
```{r data}
source(file = "rats-data.R")
rats_data
```

### Model Definition 

The following model definition is taken from [`OpenBUGS` tutorial](http://www.openbugs.net/Examples/Rats.html). 

```{r model}
# define model in strings 
rats_model <- function() {
    # likelihood 
    for( i in 1 : N ) {
        for( j in 1 : T ) {
            Y[i , j] ~ dnorm(mu[i , j],tau.c)
            mu[i , j] <- alpha[i] + beta[i] * (x[j] - xbar)
            # deleted unnecessary lines
        }
        alpha[i] ~ dnorm(alpha.c,alpha.tau)
        beta[i] ~ dnorm(beta.c,beta.tau)
    }
    
    # prior 
    tau.c ~ dgamma(0.001,0.001)
    sigma <- 1 / sqrt(tau.c)
    alpha.c ~ dnorm(0.0,1.0E-6)   
    alpha.tau ~ dgamma(0.001,0.001)
    beta.c ~ dnorm(0.0,1.0E-6)
    beta.tau ~ dgamma(0.001,0.001)
    alpha0 <- alpha.c - xbar * beta.c 
}

# write model to disk 
# model.file must be renamed with .txt rather than .bug
rats_model_file <- tempfile(pattern = "model_", fileext = ".txt") 
write.model(rats_model, con = rats_model_file)

# print file 
readLines(rats_model_file)
```

### Model Run

```{r run}
rats_fit <- bugs(
    data = rats_data,
    inits = list(
        # source: http://www.openbugs.net/Examples/Ratsinits.html
        list(
            "alpha" = rep(250, times = rats_data[["N"]]),
            "beta"  = rep(6, times = rats_data[["N"]]),
            "alpha.c" = 150,
            "beta.c"  = 10,
            "tau.c"   = 1,
            "alpha.tau" = 1,
            "beta.tau"  = 1
        )
    ),
    model.file = rats_model_file, 
    parameters.to.save = c("alpha0", "beta.c", "sigma"), 
    # MCMC parameters to be consistent with
    # http://www.openbugs.net/Examples/Rats.html
    n.chains = 1,
    n.burnin = 1000,
    n.iter = 1000 + 10000, 
    # other params 
    digits = 9, 
    bugs.seed = sample(x = seq_len(14), size = 1) # getting some randomness 
)
```

```{r summary}
print(rats_fit)
```

Comparing to the [results from OpenBUGS tutorial](http://www.openbugs.net/Examples/Rats.html): 

```{r web_result, echo=FALSE, fig.align="center"}
knitr::include_graphics("http://www.openbugs.net/Examples/rats3.bmp")
```

Note that the model did not specify constraints on `alpha` and `beta`'s hence creating *identifiability* issues. 

## Session Info

```{r session}
utils::sessionInfo()
```

